---
title: "01 ggplot"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```


## geom_col() or geom_bar() ?


```{r quiz}
quiz(
  question("You have a dataset with the countries of Europe in one column and their populations in a second column. To produce a bar chart, should you use...",
    answer("geom_bar()"),
    answer("geom_col()", correct = TRUE)
  )
)
```

## Midwest Barchart

Make a bar chart of the number of counties in each of the ve US midwest states. Use the midwest dataset from
ggplot2

```{r barchart1, exercise=TRUE, exercise.lines = 5}
midwest
```

```{r barchart1-hint}
midwest %>% 
  ggplot(aes(state))
```

## Penguin Barchart

Make a bar chart of the number of each species of pengiun from the penguins dataset in palmerpenguins. Extra marks for orienting the bars horizontally, extra-extra marks for arranging them by increasing number (try google, or the hint).

```{r barchart2, exercise=TRUE, exercise.lines = 5}
palmerpenguins::penguins
```

```{r barchart2-hint}
palmerpenguins::penguins %>% 
  ggplot(aes(fct_infreq(species))) + 
  coord_flip()
```

## Sleep Barchart

Make a bar chart of the 12 Carnivora total sleep times from the msleep dataset in ggplot2

```{r barchart3, exercise=TRUE, exercise.lines = 5}
msleep %>% 
  dplyr::filter(order == "Carnivora" ) %>% 
  mutate(name = fct_reorder(name, sleep_total))
```

```{r barchart3-hint}
msleep %>% dplyr::filter(order == "Carnivora" ) %>% 
  mutate(name = fct_reorder(name, sleep_total)) %>% 
  ggplot(aes(sleep_total))
```

