---
title: "ggplot - a Pantheon of Visualisations"
author: "Eugene"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output:
  xaringan::moon_reader:
    css: [TUD01, TUD02, tamu-fonts]
---

class: center, inverse

# Basic Picture of ggplot

- aesthetics
    - values that each individual observation (row) has
    - will be different for each observation
- atrributes
    - values that are shared between all points
    - decide to make everything mint green
- layers
    - each visualisation is built sequentially
    - add features in layers, one on top of the last
    - examples: add a plot title, change an axis scale....
    
---

```{r ggplot_example, echo = F, message = F, warning = F, cache = T}
library(tidyverse)
library(ggtext)
library(hrbrthemes)

who %>% 
  filter(iso3 == "IRL") %>% 
  select(year, new_sp_f65, new_sp_m65) %>% 
  drop_na() %>% 
  pivot_longer(cols = -year, names_to = "Gender", values_to = "Incidence") %>% 
  ggplot(aes(x = year, y = Incidence, col = Gender)) + 
  geom_point() +
  geom_smooth(se = F) + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(subtitle = "Incidence of Tuberculosis in Ireland",
       title = "<span style = 'color: red;'>Women</span> and 
       <span style = 'color: blue;'>Men</span>",
       caption = "@Data from WHO") +
  theme(plot.title = element_markdown()) +
  theme(text = element_text(family = "Ink Free", size = 24),
        axis.title = element_text(family = "Ink Free", size = 24, face = "bold"))
```


---


# Picturing Data Different Ways with ggplot

### We're going to set out some of the options for looking at data

### these depend on what kind of data you have

### and what you want to investigate

Lots of these come from [Top 50 Visualizations in R](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#5.%20Composition)



---

<ol>
<li> Visualising Amounts
<li> Visualising Proportions
<li><span style="color: red;"> Visualising Distributions </span>
<li><span style="color: red;"> Visualising Relationships </span>
<li><span style="color: red;"> Visualising Time Series </span>
<li><span style="color: red;"> Visualising Groups </span>
<li><span style="color: red;"> Visualising Networks </span>
<li><span style="color: red;"> Visualising Spatial Data </span>

Items in <span style="color: red;">red</span> we'll cover next week.

---

class: center, inverse

# Visualising Amounts

- barplot
- dot plot
- lollipop plot



---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = F)
```

```{r libraries, cache = F}
library(tidyverse)
library(scales)
library(gapminder)
library(ggrepel)
library(glue)
library(waffle)
library(treemapify)
library(ggridges)
library(lubridate)
library(boxoffice)
my_colour = "firebrick4"
ggplot2::theme_set(ggplot2::theme_minimal())
ggplot2::theme_update(text = ggplot2::element_text(family = "Ink Free", 
                                 size = 20, 
                                 colour = my_colour),
             axis.text = ggplot2::element_text(colour = my_colour),
             rect = element_rect(colour = my_colour),
             line = element_line(colour = my_colour))
caption = "@DataVis_2020 Eugene"

my_ordinal_date <- function(dates){
     dayy <- day(dates)
     suff <- case_when(dayy %in% c(11,12,13) ~ "th",
                       dayy %% 10 == 1 ~ 'st',
                       dayy %% 10 == 2 ~ 'nd',
                       dayy %% 10 == 3 ~'rd',
                       TRUE ~ "th")
     paste0(dayy, suff)
 }

```

```{r data, cache = T}
# url <- paste0("https://raw.githubusercontent.com/rfordatascience/",
# "tidytuesday/master/data/2019/2019-12-17/dog_descriptions.csv")
# dogs <- read_csv(url)
boxoffice_date <- Sys.Date()-7
movies <- boxoffice(boxoffice_date) %>% 
  mutate(gross = gross / 1e3,
         movie_name = movie,
         movie = abbreviate(movie))

```



```{r barplot, echo = T, , fig.height=6}
diamonds %>% 
  ggplot(aes(cut)) + 
  geom_bar(fill = "firebrick4", alpha = 0.6) + 
  labs(caption = "@Data tidyverse")
```

---

```{r barplot2, echo=TRUE, fig.height=5}
diamonds %>% 
  ggplot(aes(cut)) + 
  geom_bar(fill = "firebrick4") + 
  ggtitle("Proportion of Cuts of Diamonds") + 
  labs(caption = "@Data tidyverse") +
  coord_flip() +
  theme(axis.text.x = element_blank()) + 
  theme(axis.title = element_blank()) + 
  theme(title = element_text(face = "bold"))
```


---

```{r box_office, echo = T, fig.height=4}
boxoffice_date_string <- glue::glue("{my_ordinal_date(boxoffice_date)}, {month(boxoffice_date, label = T, abbr = F)}, {year(boxoffice_date)}")
movies %>% mutate(movie = fct_reorder(movie, gross)) %>% 
  slice_head(n=10) %>% 
  ggplot(aes(movie, gross)) + 
  geom_col(fill = "firebrick4") + 
  scale_y_continuous(breaks = scales::breaks_extended(8), 
                     labels = scales::label_dollar(scale = 1)) + 
  labs(title = glue::glue("Box Office for {boxoffice_date_string}"),
       caption = "@Data from BoxOfficeMojo",
       y = "Gross ($k)") +
  coord_flip()
```


---

```{r colplot_gapminder, echo = T, fig.height=6}
gapminder %>% 
  filter(year == "2007") %>% 
  group_by(continent) %>% 
  summarise(life__expectancy_2007 = mean(lifeExp)) %>%
  ggplot(aes(continent, life__expectancy_2007)) + 
  geom_col() + 
  coord_flip() +
  labs(caption = "@Data Gapminder (WHO)")
```

---

```{r dotplot_gapminder, echo = T, fig.height=6}
gapminder %>% 
  filter(year == "2007") %>% 
  group_by(continent) %>% 
  summarise(life__expectancy_2007 = mean(lifeExp)) %>%
  ggplot(aes(continent, life__expectancy_2007)) + 
  geom_point(colour = "firebrick4", size = 4) + 
  coord_flip() +
  labs(caption = "@Data Gapminder (WHO)")
```

---

```{r lollypop_gapminder, echo = T, fig.height=5}
gapminder %>% 
  filter(year == "2007") %>% 
  group_by(continent) %>% 
  summarise(life_expectancy_2007 = mean(lifeExp)) %>%
  mutate(continent = fct_reorder(continent, life_expectancy_2007)) %>% 
  ggplot(aes(continent, life_expectancy_2007)) + 
  geom_point(colour = "firebrick4", size = 10) + 
  geom_segment(aes(x = continent, xend = continent, y = 0, yend = life_expectancy_2007),
               size = 2, colour = "firebrick4") +
  coord_flip() +
  labs(caption = "@Data Gapminder (WHO)")
```

---

## Summary of Amounts

- have categorical variable and a count / quantity
- geom_bar() does a count, does some processing for you
- geom_col() when you have quantities already
- dot plot with geom_point() doesn't have to include origin
- lollypop pop same as geom_col() but with less ink. Uses geom_point() and geom_segment()
- coord_flip() useful, especially when text labels are long
- important to get order of categorical variable right, fct_reorder() useful

---

class: center, inverse

# Visualising Proportions

- stacked barplot
- pie chart
- waffle plot
 
---

```{r pop_gapminder, echo = T, fig.height=5}
gapminder %>%
  filter(year == "2007") %>%
  mutate(pop = round(pop / 1e6, 0)) %>%
  group_by(continent) %>%
  summarise(population = sum(pop)) %>%
  mutate(continent = fct_reorder(continent, population)) %>%
  ggplot(aes(1, population, fill = continent)) +
    geom_col(position = "stack", show.legend = F) +
    ylab("Population (Millions)") + 
    geom_text_repel(aes(label = glue("{continent}: {population}M")),  
              position = position_stack(vjust = .5), size = 6) +
    theme_void()
```

---


```{r pie_gapminder, echo = T, fig.height=4}
gapminder %>%
  filter(year == "2007") %>%
  mutate(pop = round(pop / 1e6, 0)) %>%
  group_by(continent) %>%
  summarise(population = sum(pop)) %>%
  mutate(continent = fct_reorder(continent, population)) %>%
  ggplot(aes(1, population, fill = continent)) +
    geom_col(position = "stack", show.legend = F) +
    ylab("Population (Millions)") + 
    geom_text_repel(aes(label = glue("{continent}: {population}M")),  
              position = position_stack(vjust = .5), size = 6) +
    theme_void() +
    coord_polar(theta = "y")
```

---

```{r waffle, echo=T, fig.height=6}
gapminder %>%
  filter(year == "2007") %>%
  mutate(pop = pop / 1e3) %>%
  group_by(continent) %>%
  summarise(population = (sum(pop) / 1e5) %>% ceiling()) %>%
  mutate(continent = fct_reorder(continent, population)) %>% pull(population) %>%
  waffle(rows = 5, legend_pos = "none")
```

---

```{r treemap, echo = T, fig.height=5, cache = T}
read_delim(glue("https://raw.githubusercontent.com/eugene100hickey/DataVis_2020/master/ireland_cattle.csv?token=ADCNXCLOQWXRR3HFVCW5JRK7QNTUW"), delim = "\t") %>% 
  ggplot(aes(area = herd, fill = province, label = county)) + 
    geom_treemap() + 
    geom_treemap_text(fontface = "italic", 
                      colour = "white", 
                      place = "centre", 
                      grow = T) +
  labs(caption = "@Data from eurostat and IBGF")
```

---

## Summary of Proportions

- have categorical variable and a count / quantity
- want to know relative amounts of each
- options are stacked bars, pie, waffle, treemap
- I'm not a huge fan of any of these