---
title: "ggplot - a Pantheon of Visualisations"
author: "Eugene"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output:
  xaringan::moon_reader:
    css: [TUD01, TUD02, tamu-fonts]

---


# Picturing Data Different Ways with ggplot

### We're going to set out some of the options for looking at data

### these depend on what kind of data you have

### and what you want to investigate

Lots of these come from [Top 50 Visualizations in R](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#5.%20Composition)



---

<ol>
<li> Visualising Amounts
<li> Visualising Proportions
<li><span style="color: red;"> Visualising Distributions </span>
<li><span style="color: red;"> Visualising Relationships </span>
<li>Visualising Time Series
<li>Visualising Groups
<li>Visualising Networks
<li>Visualising Spatial Data

Items in <span style="color: red;">red</span> we'll cover this week.


---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, cache = F)
```

```{r libraries, cache = F}
library(tidyverse)
library(scales)
library(gapminder)
library(ggrepel)
library(glue)
library(waffle)
library(treemapify)
library(ggridges)
library(lubridate)
library(boxoffice)
my_colour = "firebrick4"
ggplot2::theme_set(ggplot2::theme_minimal())
ggplot2::theme_update(text = ggplot2::element_text(family = "Ink Free", 
                                 size = 20, 
                                 colour = my_colour),
             axis.text = ggplot2::element_text(colour = my_colour),
             rect = element_rect(colour = my_colour),
             line = element_line(colour = my_colour))
caption = "@DataVis_2020 Eugene"

my_ordinal_date <- function(dates){
     dayy <- day(dates)
     suff <- case_when(dayy %in% c(11,12,13) ~ "th",
                       dayy %% 10 == 1 ~ 'st',
                       dayy %% 10 == 2 ~ 'nd',
                       dayy %% 10 == 3 ~'rd',
                       TRUE ~ "th")
     paste0(dayy, suff)
 }

```

```{r data, cache = T}
# url <- paste0("https://raw.githubusercontent.com/rfordatascience/",
# "tidytuesday/master/data/2019/2019-12-17/dog_descriptions.csv")
# dogs <- read_csv(url)
boxoffice_date <- Sys.Date()-7
movies <- boxoffice(boxoffice_date) %>% 
  mutate(gross = gross / 1e6,
         movie = abbreviate(movie))

```



class: center, inverse

# Visualising Distributions

- histograms
- density plots
- boxplot
- violin plot
- ridge plots

---

```{r histogram, echo = T, fig.height=5}
basketball <- read_csv(glue("https://raw.githubusercontent.com/eugene100hickey/DataVis_2020/master/basketball.csv?token=ADCNXCIJQV7TBVNWLF77UQC7RGRA6"))
basketball %>% 
  ggplot(aes(weight)) + 
  geom_histogram(fill = "firebrick4", bins = 50) + 
  labs(caption = "@Data from Kaggle", x = "weight (kg)", y = "",
       title = "Weight of NBA Players")
```

---

```{r histogram_dodge, echo = T, fig.height=6}
basketball %>% 
  ggplot(aes(weight, fill = position)) + 
  geom_histogram(bins = 20, position = "dodge") + 
  labs(caption = "@Data from Kaggle", x = "weight (kg)", y = "",
       title = "Weight of NBA Players by Position")
```

---

```{r density, echo = T, fig.height=6}
basketball %>% 
  ggplot(aes(weight, col = position)) + 
  stat_density(geom = "line", position = "identity") + 
  labs(caption = "@Data from Kaggle", x = "weight (kg)", y = "",
       title = "Weight of NBA Players by Position") +
  geom_rug()
```

---


```{r boxplots, echo = T, fig.height=6}
basketball %>% 
   ggplot(aes(x = position, y = weight, colour = position)) + 
   geom_boxplot(show.legend = F) + 
   labs(caption = "@Data from Kaggle", y = "weight (kg)", x = "position",
        title = "Weight of NBA Players by Position")
```

---

```{r boxplots_points, echo = T, fig.height=6}
basketball %>% 
   ggplot(aes(x = position, y = weight, colour = position)) + 
   geom_boxplot(show.legend = F) + 
   labs(caption = "@Data from Kaggle", y = "weight (kg)", x = "position",
        title = "Weight of NBA Players by Position") +
  geom_jitter(size = 0.4, alpha = 0.2, show.legend = F)
```

---

```{r violins, echo = T, fig.height=6}
basketball %>% 
   ggplot(aes(x = position, y = weight, colour = position)) + 
   geom_violin(show.legend = F) + 
   labs(caption = "@Data from Kaggle", y = "weight (kg)", x = "position",
        title = "Weight of NBA Players by Position") +
   geom_jitter(size = 0.4, alpha = 0.2, show.legend = F)
```

---
```{r ridges, echo = T, fig.height=6}
gapminder %>% ggplot(aes(x = lifeExp, y = factor(year))) +
  geom_density_ridges(fill = "firebrick4", colour = "firebrick4", alpha = 0.4) +
  theme_ridges() +
  labs(x = "Life Expectancy (years)", y = "", caption = "@Data Gapminder (WHO)")
```

---

## Summary of Distributions

- hugely important
- great way to explore your data / introduce it to others
- make sure you show you data when possible
    - use *geom_rug()*
    - use *geom_jitter()*
    - if lots of points, then use *alpha* to mute them
    
---

class: center, inverse

# Visualising Relationships

- scatter plots
    - encircling
    - jittering
    - using colour / size / shape 
    - fitting lines
    - histograms and boxplots on the axes (and geom_rug())
- line plots
- correlation

---

```{r scatterplot, echo = T, fig.height=6}
library(dslabs)

stars %>% 
  ggplot(aes(temp, magnitude, col = type)) + 
  geom_point(show.legend = F)
```


---

```{r encircleplot, echo = T, fig.height=4}
library(ggalt)

stars %>% 
  ggplot(aes(temp, magnitude, col = type)) + 
  geom_point(show.legend = F) +
  geom_encircle(data = stars %>% 
                  filter(type == "B" | (type == "M" & magnitude > 9)), show.legend = F) +
  scale_x_log10() + 
  theme_minimal() + 
  annotate("text", x = c(15000, 5000), y = c(-4, 14), 
           label = c("Type B Stars", "Faint Type M Stars"), 
           col = c("blue", "olivedrab3"), family = "Ink Free", size = 4, fontface = 2) + 
  scale_color_viridis_d()
```

---

```{r jitter, echo = T, fig.height=6, fig.width=12}
library(HistData)
library(patchwork)

scatter <- Galton %>% ggplot(aes(parent, child)) + geom_point()
jittered <- Galton %>% ggplot(aes(parent, child)) + geom_jitter(width = 0.4, height = 0.4)

scatter + plot_spacer() + jittered 
```

