---
title: "Mapping in R"
author: "Eugene"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output:
  xaringan::moon_reader:
    css: [TUD01, TUD02, tamu-fonts]

header-includes:
- \usepackage{multicol}
- \newcommand{\btwocol}{\begin{multicols}{2}}
- \newcommand{\etwocol}{\end{multicols}}
---

class: center, inverse

# Maps
<br>

### Vibrant area of development in R
### Large, and growing, number of packages available
### And, both fun and fascinating
    
---

## We'll begin by looking at some maps, to get a flavour of what R can do

## Then we'll discuss how it does it

---

- drawing maps from google, different map types
    - maps of Dublin in terrain, satellite, and roadmap 
- adding layers to our maps
    - Irish small airport locations
- doing some statistics on our maps
    - where are Irish schools concentrated?
    - incidence of violence in counties of Sao Paulo
- using facets in maps
    - GDP in European countries
- interactive maps
    - meteor locations
- world projections


---

# Data with Location Information

```{r setup, echo = F, message=F, warning=F}
library(tidyverse)
library(knitr)
library(kableExtra)
library(sf)
library(rworldmap)
library(tmap)
library(leaflet)
library(ggmap)
library(ggrepel)
library(ggsn)
library(geofacet)
library(rvest)
library(stringi)
require(handlr)
library(lubridate)
library(colorspace)
```

```{r echo = T, message=F, warning=F, cache=TRUE, results = 'asis'}
schools <- read_csv("http://airo.maynoothuniversity.ie/files/dDATASTORE/education/csv/primary_schools_2013_2014.csv")
schools %>% 
  select(Off_Name, Add_1, T_13_14, Long, Lat) %>% 
  head(10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

---

# Data with Location Information

```{r echo = T, message=F, warning=F, cache=TRUE}
covid <- read_csv("/home/eugene/Desktop/Academic/Research/CoVid19/time_series_covid_19_confirmed.csv")
covid %>% 
  select("Country/Region", Lat, Long, "3/12/20", "3/13/20", "3/14/20") %>% 
  head(10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```
---

# Data with Location Information

```{r echo = T, message=F, warning=F, cache=TRUE}
airports <- read_csv("http://data-osi.opendata.arcgis.com/datasets/1ace692fe9fd4a648ee78347ad571de2_22.csv")
airports %>% 
  select(X, Y, Airport = NAMN1) %>% 
  head(10) %>% 
  kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## Some Examples of Maps

```{r map1, warning = F, message=F, cache = T, fig.height=5, fig.width=5, fig.align='center'}
p <- ggmap(get_googlemap(center = c(lon = -7.61, lat = 53.463),
                     zoom = 7, scale = 2,
                     maptype ='terrain',
                     color = 'color'))
p
```

---

## Some Examples of Maps

```{r map2, warning = F, message=F, cache = T, fig.height=5, fig.width=5, fig.align='center'}
p <- ggmap(get_googlemap(center = c(lon = -6.32, lat = 53.3),
                     zoom = 11, scale = 2,
                     maptype ='satellite',
                     color = 'color'))
p
```

---

## Some Examples of Maps

```{r map3, warning = F, message=F, cache = T, fig.height=5, fig.width=5, fig.align='center'}
ggmap(get_googlemap(center = c(lon = -6.32, lat = 53.3),
                     zoom = 11, scale = 2,
                     maptype ='roadmap',
                     color = 'color'))
```

---

## Maps with Added Layers


```{r map4, warning = F, message=F, cache = T, fig.height=5, fig.width=5, fig.align='center'}
ggmap(get_stamenmap(bbox = c(top = 55.42, bottom = 51.38, right = -5.06, left = -10.7), zoom = 7, scale = 2, maptype ='watercolor', color = 'color')) + 
  geom_point(data = airports, size = 2,
             aes(x = X, y = Y), col = "firebrick4") + 
  geom_text_repel(data= airports, aes(label = NAMN1, x = X, y = Y), col = "firebrick4")
```

---
```{r map5, warning = F, message=F, cache = T, fig.height=5, fig.width=5, fig.align='center'}
ggmap(get_googlemap(center = c(lon = -7.833, lat = 53.333),
                     zoom = 7, scale = 2,
                     maptype ='roadmap',
                     color = 'bw')) +
  stat_density2d(aes(x = Long, y = Lat, fill = ..level..),
                 alpha = 0.1, size = 0.01, bins = 50, 
                 data = schools, geom = "polygon") + 
  guides(fill = F) +
  labs(title = "Where are Schools in Ireland", caption = "@Data from NUIM")

```



---

```{r map6, warning = F, message=F, cache = T, fig.height=5, fig.width=5, fig.align='center'}
data("domestic_violence")
ggplot(domestic_violence, aes(fill = Scaled)) +
    geom_sf() +
    scale_fill_continuous(low = "#fff7ec", high = "#7F0000") + 
    blank() +
    north(domestic_violence) +
    scalebar(domestic_violence, dist = 4, dist_unit = "km",
             transform = TRUE, model = "WGS84")
```

---

## Maps Layout of Plots - geofacet

```{r geofacet, message=F, warning=F, echo=T, cache=T, fig.height=5}
load("/home/eugene/Downloads/eu_gdp.rda")
ggplot(eu_gdp, aes(year, gdp_pc)) +
  geom_line(color = "steelblue") +
  facet_geo(~ name, grid = "eu_grid1", scales = "free_y") +
  scale_x_continuous(labels = function(x) paste0("'", substr(x, 3, 4))) +
  ylab("GDP Per Capita in Relation to EU Index (100)") +
  theme_bw()
```

---

## Maps Layout of Plots - geofacet

```{r ireland_grid, echo = F, message=F, warning=F, cache=TRUE}
url <- "https://en.wikipedia.org/wiki/List_of_Irish_counties_by_population"
w <- read_html(url)
pathway_data_html <- html_nodes(w, ".headerSort , td")
ev <- html_text(pathway_data_html)
ev <- ev %>% 
  gsub(pattern = "%", replacement = "") %>% 
  gsub(pattern = ",", replacement = "") %>% 
  gsub(pattern = "\\\n", replacement = "") %>% 
  gsub(pattern = "–", replacement = "-")
number <- ev[c(T, F, F, F, F, F)]
county <- ev[c(F, T, F, F, F, F)]
population <- ev[c(F, F, T, F, F, F)]
density <- ev[c(F, F, F, T, F, F)]
province <- ev[c(F, F, F, F, T, F)]
change <- ev[c(F, F, F, F, F, T)]
irish_counties <- data_frame(county = county, 
                  population = as.numeric(population), 
                  province = province, 
                  density = as.numeric(density), 
                  change = as.numeric(change))

irish_counties <- irish_counties[-c(5, 6, 10),]
irish_counties$county[6] <- "Derry"

ireland_grid <- read_csv("/home/eugene/Desktop/Academic/Research/R_Fun/Irish_counties/ireland.csv")

irish_counties %>% ggplot(aes(1, change, fill = province)) +
  geom_col() +
  facet_geo(~county, grid = ireland_grid) +
  theme(axis.text.x = element_blank(),
        panel.background = element_rect(fill = "darkgreen"),
        panel.grid = element_blank()) +
  scale_fill_manual(values = c("lightgreen", "blue", "red", "white")) +
  guides(fill = F) +
  labs(title = "Population Growth in Ireland by County",
       caption = "@Data wikipedia")
```


---

## Interactive Maps with leaflet()

```{r leaflet, warning = F, message=F, echo=F, fig.align='center', fig.height = 6, fig.width = 10, cache = F, eval=T}
meteors <- read_csv("/home/eugene/Desktop/Academic/Research/Blogs/Meteorites/Meteorite_Landings.csv")

names(meteors)[5] <- "mass_g"

meteors <- meteors %>% 
  mutate(year=year(dmy_hms(year))) %>% 
  mutate(kg=mass_g/1000) %>% 
  na.omit %>% 
  filter(kg > 10)

pal <- colorNumeric("Reds", 
                    domain=(meteors %>% 
                              filter(year>1780))$year)

magnification <- 10000

map <- meteors %>% 
  filter(year > 1800) %>% 
  leaflet() %>% 
  addProviderTiles(provider = "Esri") %>% 
  setView(lng=-5, lat=50, zoom=2) 

map %>% 
  addCircles(lat=~reclat,
             lng=~reclong,
             popup = ~paste(name,"<br>",  
                            "Mass", kg, "kg", "<br>",
                            "Year", year),
             color = ~pal(year),
             radius=~(((kg)^0.25)*magnification))
```

---

## Cartograms

- Distort Size (and Shape) to Correspond to some Variable
- Map of US income where each state has an area proportional to its population

```{r cartogram, echo=F, fig.align='center', fig.height = 6, fig.width = 6, cache = T}
load("/home/eugene/Downloads/US_income.rda")
load("/home/eugene/Downloads/US_income_cartogram.rda")

US_income <- mutate(
  US_income,
  income_bins = cut(
      ifelse(is.na(median_income), 25000, median_income), # hide missing value
      breaks = c(0, 40000, 50000, 60000, 70000, 80000),
      labels = c("< $40k", "$40k to $50k", "$50k to $60k", "$60k to $70k", "> $70k"),
      right = FALSE
    )
  )
US_income_cartogram$income_bins <- US_income$income_bins

US_income_cartogram$income_bins <- US_income$income_bins
p <- ggplot(US_income_cartogram, aes(fill = income_bins)) + 
  geom_sf(color = "grey30", size = 0.2) + coord_sf(datum = NA, expand = FALSE) +
  scale_x_continuous(limits = c(-3900000, 2500000)) +
  scale_fill_discrete_sequential(
    h1 = -83, h2 = 20, c1 = 30, cmax = 40, c2 = 0, l1 = 20, l2 = 100, p1 = 1, p2 = 1.2, rev = TRUE, 
    name = "median income",
    nmax = 7,
    order = 2:6,
    guide = guide_legend(
      override.aes = list(colour = "white", size = 1),
      reverse = TRUE
    )
  ) +
  theme(
    plot.background = element_rect(fill = "cornsilk"),
    legend.position = c(0, 1),
    legend.justification = c(0, 1),
    legend.spacing.x = grid::unit(3, "pt"),
    legend.spacing.y = grid::unit(3, "pt"),
    legend.title = element_text(hjust = 0.5),
    legend.key.width = grid::unit(18, "pt"),
    legend.key.height = grid::unit(15, "pt"),
    plot.margin = margin(3, 3, 3, 1.5)
  )
p

```


---

## Mercator
```{r mercator, echo=F, fig.align='center', fig.height = 6, fig.width = 6, cache = T}
world_sf <- sf::st_as_sf(rworldmap::getMap(resolution = "low"))

crs_longlat <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
crs_mercator <- "+proj=merc"
# calculate bounding box in transformed coordinates
mercator_bbox <- 
  rbind(c(-180, -85), c(180, 85)) %>%
  st_multipoint() %>%
  st_sfc(
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>% 
  st_transform(crs = crs_mercator)
ggplot(world_sf) + 
  geom_sf(fill = "#E69F00B0", color = "black", size = 0.5/.pt) + 
  scale_x_continuous(name = "longitude", breaks = seq(-120, 120, by = 60)) +
  scale_y_continuous(name = "latitude", breaks = seq(-80, 80, by = 20)) +
  coord_sf(
    xlim = mercator_bbox[[1]][, 1],
    ylim = mercator_bbox[[1]][, 2],
    expand = FALSE,
    crs = crs_mercator
  ) + 
  theme(
    panel.background = element_rect(fill = "#56B4E950", color = "#56B4E950"),
    panel.grid.major = element_line(color = "gray30", size = 0.25),
    axis.ticks = element_line(color = "gray30", size = 0.5/.pt)
  )

```

---

## Mollweide

```{r mollweide, echo=F, fig.align='center', fig.height = 10, fig.width = 10, cache = T}
world_mol <- st_transform(world_sf, crs = "+proj=moll")

mol_bbox <- 
  rbind(c(-180, -85), c(180, 85)) %>%
  st_multipoint() %>%
  st_sfc(
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>% 
  st_transform(crs = "+proj=moll")
```

```{r mercator_plot, message=F, warning=F, echo=F, cache=T}
ggplot(world_mol) + 
  geom_sf(fill = "#E69F00B0", color = "black", size = 0.5/.pt) + 
  geom_sf(data = mol_bbox, fill = NA, color = "grey30", size = 0.5/.pt) +
  scale_x_continuous(name = "longitude", breaks = seq(-120, 120, by = 60)) +
  scale_y_continuous(name = "latitude", breaks = seq(-80, 80, by = 20)) +
  theme(
    panel.background = element_rect(fill = "#56B4E950", color = "#56B4E950"),
    panel.grid.major = element_line(color = "gray30", size = 0.25),
    axis.ticks = element_line(color = "gray30", size = 0.5/.pt)
  )
```
---

##Interrupted Goode Homolosine

```{r interrupted_goode_homolosine, echo=F, fig.align='center', fig.height = 10, fig.width = 10, cache = T}
## Interrupted Goode homolosine
crs_goode <- "+proj=igh"
# projection outline in long-lat coordinates
lats <- c(
  90:-90, # right side down
  -90:0, 0:-90, # third cut bottom
  -90:0, 0:-90, # second cut bottom
  -90:0, 0:-90, # first cut bottom
  -90:90, # left side up
  90:0, 0:90, # cut top
  90 # close
)
longs <- c(
  rep(180, 181), # right side down
  rep(c(80.01, 79.99), each = 91), # third cut bottom
  rep(c(-19.99, -20.01), each = 91), # second cut bottom
  rep(c(-99.99, -100.01), each = 91), # first cut bottom
  rep(-180, 181), # left side up
  rep(c(-40.01, -39.99), each = 91), # cut top
  180 # close
)
goode_outline <- 
  list(cbind(longs, lats)) %>%
  st_polygon() %>%
  st_sfc(
    crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
  ) %>% 
  st_transform(crs = crs_goode)
# bounding box in transformed coordinates
xlim_goode <- c(-21945470, 21963330)
ylim_goode <- c(-9538022, 9266738)
goode_bbox <- 
  list(
    cbind(
      c(xlim_goode[1], xlim_goode[2], xlim_goode[2], xlim_goode[1], xlim_goode[1]), 
      c(ylim_goode[1], ylim_goode[1], ylim_goode[2], ylim_goode[2], ylim_goode[1])
    )
  ) %>%
  st_polygon() %>%
  st_sfc(crs = crs_goode)
# area outside the earth outline
goode_without <- st_difference(goode_bbox, goode_outline)
```

```{r goode_plot, message=F, warning=F, echo=F, cache=T}
ggplot(world_sf) + 
  geom_sf(fill = "#E69F00B0", color = "black", size = 0.5/.pt) +
  geom_sf(data = goode_without, fill = "white", color = NA) +
  geom_sf(data = goode_outline, fill = NA, color = "grey30", size = 0.5/.pt) +
  scale_x_continuous(name = NULL, breaks = seq(-160, 160, by = 20)) +
  scale_y_continuous(name = NULL, breaks = seq(-80, 80, by = 20)) + 
  coord_sf(xlim = 0.95*xlim_goode, ylim = 0.95*ylim_goode, expand = FALSE, crs = crs_goode, ndiscr = 1000) + 
  theme(
    panel.background = element_rect(fill = "#56B4E950", color = "white", size = 1),
    panel.grid.major = element_line(color = "gray30", size = 0.25),
    plot.margin = margin(1.5, 1.5, 24, 1.5)
  )
```


---

## Packages for Drawing Maps

- lots of packages ('cus lots of people interested)
- *ggmap*
    - useful for drawing maps with layers
- *leaflet*
    -useful for interactive maps
- *tmap*
    - useful for both, this is the one to use
    - see [vignette](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html) for instructions
    
---

## Types of Maps

- Two main categories
    - shapefiles
    - rasters
- shapefiles are composed of geometical objects of points, lines or polygons
- rasters are arrays of pixels essentially
- we've used shapefiles here
    - new format called *sf* replacing *sp*
    
---

## Some Libraries for Maps

- the ones mentioned above, *ggmap*, *leaflet*, and *tmap*
- *sf* for layers like *geom_sf()*
- *sp* still useful for legacy reasons
- *spData* has lots of shapefiles
- *rworldmap*, likewise

---

## Some Libraries for Maps

- *rnaturalearth*
- *rnoaa* for weather data
- *RgoogleMaps* is the gateway to all the maps.google stuff
    - but you have to register now to get all this
- *usmap*, kind of what it suggests
- *albersusa*, shapefile projection of the US
- can also google __shapefile__ plus some area to see what you get
- *osmdata* gives access to open street map data
    
---

## More Involved Packages

- *RQGIS* is an R interface for a _Graphical Information System_
- *spatstat* is for spatial statistics
    - clustering, correlations, point pattern analysis, kriging, etc
    - one of the largest libraries in all of R
- *RSAGA* for geoprocessing and terrain analysis
- *SPADAR* for astronomical mapping


---
## Some Nice Add on Libraries

- *ggsn*, we used this to get a scale bar and compass for Sao Paolo
- *ggspatial* likewise
- *geofacet* for the European GDP plot
- *ggrepel* for nice labelling in plots

---

## Some Resources

- book by [Wilke](https://serialmentor.com/dataviz/) has a chapter on maps
    - several of maps above taken from [here](https://github.com/clauswilke/dataviz/blob/master/geospatial_data.Rmd)
- article by [littlemissdata](https://www.littlemissdata.com/blog/ggmap-updated) uses *ggmap*
    - also includes instructions to register to access *get_googlemap()*
- online book on [Geocomputation with R](https://geocompr.robinlovelace.net/)
- list of [Coordinate Reference Systems](https://spatialreference.org/)
- article on using [Open Street Map Data](https://dominicroye.github.io/en/2018/accessing-openstreetmap-data-with-r/)
- tutorial from the [University of Chicago](https://spatialanalysis.github.io/)

---

```{r my_code, ref.label=knitr::all_labels(), eval=F}

```


