---
title: "Flipbooking"
subtitle: "With flipbookr and xaringan"
author: "Gina Reynolds, December 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
library(ggthemes)
library(paletteer)
```



## `break_type = "rotate"`

And break_type = "rotate" is used to to cycle through distinct lines of code. The special comment to indicate which lines should be cycled through is *#ROTATE*. 

--

Sometimes, you will need to set the chunk option to eval = F, echo = F instead of include = F.  This will be true when the input code itself cannot be executed as code to be cycled through is incompatible and would throw an error. 


---



`r chunk_reveal("my_rotate", break_type = "rotate", widths = c(50,50))`

```{r my_rotate, eval = F, echo = F}
dslabs::gapminder %>%
  group_by(continent, year) %>%
  summarise(average_fertility =
              mean(fertility, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(aes(x = year,
             y = average_fertility,
             col = continent)) +
  geom_line(size = 2) +
  scale_color_brewer(type = 'qual') + #ROTATE
  scale_color_brewer(type = 'div') + #ROTATE
  scale_color_viridis_d(option = 'magma') + #ROTATE
  scale_color_paletteer_d('rtist::warhol') + #ROTATE
  scale_color_paletteer_d('MapPalettes::irish_flag') + #ROTATE
  scale_color_paletteer_d('yarrr::southpark') + #ROTATE
  scale_color_paletteer_d('dutchmasters::pearl_earring') + #ROTATE
  scale_color_paletteer_d('dutchmasters::view_of_Delft') + #ROTATE
  scale_color_paletteer_d('ghibli::KikiLight') + #ROTATE
  scale_color_paletteer_d('LaCroixColoR::Coconut') + #ROTATE
  scale_color_paletteer_d('LaCroixColoR::PassionFruit') + #ROTATE
  scale_color_paletteer_d('ggsci::lanonc_lancet') + #ROTATE
  scale_color_paletteer_d('ggsci::default_uchicago') + #ROTATE
  scale_color_paletteer_d('ggsci::uniform_startrek') #ROTATE
```
