---
title: "Distributions"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F, fig.width = 5)
```


## Histograms


Make a histogram of the county areas (variable, *area*) from the midwest dataset from ggplot2.

```{r county, exercise=TRUE, exercise.eval=TRUE}

```

```{r county-hint}
midwest %>% 
  ggplot(aes(area)) + 
  geom_histogram()
```

## Density Plot

Make a density plot of the county areas (variable, *area*) from the midwest dataset from ggplot2.

```{r}
theme_set(theme(legend.position = "bottom"))
```


```{r density, exercise=TRUE, exercise.eval=TRUE}

```

```{r density-hint}
midwest %>% 
  ggplot(aes(area)) + 
  stat_density(geom = "line", position = "identity")
```

## Density by State

Make a density plot of the county areas from the midwest dataset from ggplot2 which has five different lines for each of the five states, (variable, *state*)

```{r density5, exercise=TRUE, exercise.eval=TRUE}

```

```{r density5-hint}
midwest %>% 
  ggplot(aes(area, col = state)) + 
  stat_density(geom = "line", position = "identity")
```

## Another Histogram

See if you can reproduce the plot using the *midwest* dataset show below. Hint, *scale_x_log10()* will be useful

```{r}
midwest %>% 
  ggplot(aes(poptotal)) + 
  geom_histogram() + 
  scale_x_log10(labels = scales::comma_format())
```


```{r filter, exercise=TRUE}

```

```{r filter-hint-1}
midwest %>% 
  ggplot(aes(poptotal)) + 
  geom_histogram()
```

```{r filter-hint-2}
midwest %>% 
  ggplot(aes(poptotal)) + 
  geom_histogram() + 
  scale_x_log10()
```

```{r filter-hint-3}
midwest %>% 
  ggplot(aes(poptotal)) + 
  geom_histogram() + 
  scale_x_log10(labels = scales::comma_format())
```
