---
title: "Fonts"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(extrafont)
library(showtext)
library(HistData)
knitr::opts_chunk$set(echo = FALSE)
```


## Google Fonts

### Categories 

*Fonts are categorised based on their style: Serif, Sans Serif, Handwriting, Display, and Monospace.*

```{r google-fonts}
quiz(
  question("Go to fonts.google.com, which of the following are Display fonts?",
    answer("Bubblegum Sans", correct = TRUE),
    answer("Raleway"),
    answer("Rye", correct = TRUE),
    answer("Love Ya Like A Sister", correct = TRUE)
  )
)
```


## Using Fonts

Use *font_add_google()* from the *showtext* package to add the *Caveat* font. Make it available with *showtext_auto()*. Use this font then to change the text style in the plot below (you'll need to add the line, *theme(text = element_text(family = "Caveat"))*) to the end of the code.

```{r cholera, exercise=TRUE, exercise.lines = 5, warning=F, message = F, fig.width=5}
Cholera %>% ggplot(aes(popn, cholera_deaths, colour = water)) +
  geom_point() +
  labs(x = "Population",
       y = "No of Cholera Deaths in 1849")
```


```{r cholera-hint, warning=F, message = F, fig.width=5}
library(showtext)
font_add_google(name = "Caveat", family = "Caveat")
showtext_auto()

Cholera %>% ggplot(aes(popn, cholera_deaths, colour = water)) +
  geom_point() +
  labs(x = "Population",
       y = "No of Cholera Deaths in 1849") +
  theme(text = element_text(family = "Caveat",
                            size = 32))
```
