---
title: "Themes"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggthemes)
library(datasets)

knitr::opts_chunk$set(echo = FALSE)
```


## Theme Elements

### panel.background 

Use *theme_get()* to find the default theme for ggplot. Now look for the *panel.background*, then *fill*, then *colour*. 

```{r panel-background, exercise=TRUE}

```

```{r panel-background-hint, warning=F, message = F, fig.width=5}
theme_get()$panel.background$fill
```

### Change theme

Now use *theme_set()* to change the theme to *theme_solarized()* from the *ggthemes* package. Find the new value for the background fill colour.


```{r solarized-background, exercise=TRUE}

```

```{r solarized-background-hint-1, warning=F, message = F, fig.width=5}
library(ggthemes)
theme_set(theme_solarized())
```


```{r solarized-background-hint-2, warning=F, message = F, fig.width=5}
library(ggthemes)
theme_set(theme_solarized())
theme_get()$panel.background$fill
```

## Geyser Plots

### Theme Change

Using the *faithful* dataset from *datasets*, make a plot of *eruptions* versus *waiting* using *theme_solarized()* from the ggthemes package


```{r faithful1, exercise=TRUE, exercise.eval=TRUE}

```

```{r faithful1-hint}
faithful %>% 
  ggplot(aes(waiting, eruptions)) + 
  geom_point() + 
  theme_solarized()
```

### Text Size Change

Now, take the plot above and change the text size to *size = 20*

```{r faithful2, exercise=TRUE, exercise.eval=TRUE}

```

```{r faithful2-hint}
faithful %>% 
  ggplot(aes(waiting, eruptions)) + 
  geom_point() + 
  theme_solarized() +
  theme(text = element_text(size = 20))
```

### Now take the plot above and remove the ticks from the x and y axes

```{r faithful3, exercise=TRUE, exercise.eval=TRUE}

```

```{r faithful3-hint}
faithful %>% 
  ggplot(aes(waiting, eruptions)) + 
  geom_point() + 
  theme_solarized() +
  theme(text = element_text(size = 20),
        axis.ticks = element_blank())
```
