---
title: "Colour Codes"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(gplots)
library(viridis)
library(RColorBrewer)
library(scales)
```

```{r samples, echo = F}
my_colours <- sample(colors(), 6)
```

## hex codes

Use the *gplot::col2hex()* and *gplot::col2rgb()* functions to find the hex codes and rgb codes for the following colours: `r my_colours`

```{r hexcode1, exercise=TRUE, exercise.lines = 5}

```

```{r hexcode2}
my_colour1 <- sample(colors(), 4)
my_colour_hex <- col2hex(my_colour1)
my_string <- glue::glue("What is the hex code for the colour, {my_colour1[3]}?")
quiz(
  question(my_string,
    answer(my_colour_hex[1]),
    answer(my_colour_hex[2]),
    answer(my_colour_hex[3], correct = TRUE),
    answer(my_colour_hex[4])
  ),
  question("What is the rgb code for the colour, cornflowerblue?",
    answer("(100, 149, 237)", correct = TRUE),
    answer("(255, 248, 220)"),
    answer("(153, 50, 204)"),
    answer("(0, 191, 255)")
  )
)
```


## Display Colours, *show_col()*

Display nine colours from the *magma()* palette (from *viridis* package) using *show_col()*

```{r viridis, exercise=TRUE, exercise.eval=TRUE}

```

```{r viridis-hint}
viridis::magma(n=9) %>% scales::show_col()
```

Type *brewer.pal.info* to get the names and types of the brewer palettes. Pick a diverging palette and display six colours using *scales::show_col()* [hint, use brewer.pal(name = "????"n = 12) where "????" is a palette name (see ?RColorBrewer::brewer.pal)]

```{r brewer, exercise=TRUE, exercise.eval=TRUE}

```

```{r brewer-hint}
RColorBrewer::brewer.pal(name = "RdYlBu", n=9) %>% scales::show_col()
```
