---
title: "Colour Plots"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE, message=F, warning=F)
library(tidyverse)
library(palmerpenguins)
library(datasets)
library(gapminder)
library(scales)
```


## Penguin Plot

Produce a ggplot from the penguins dataset from the *palmerpenguins* package. Place *bill_length_mm* on the x-axis, *bill_depth_mm* on the y, and the points are coloured by *species*. Choose the colours using *scale_color_manual()*. Pick nice colours using the chart from Stowers (or from coolor.co). Something Antarcticky would be appropriate.

```{r penguins, exercise=TRUE, exercise.eval = TRUE}

```


```{r penguins-hint}
palmerpenguins::penguins %>% 
  ggplot(aes(bill_length_mm, bill_depth_mm, colour = species)) + 
  geom_point() + 
  scale_colour_manual(values = c("#BBBBEE", "#225588", "#050511")) +
  theme_minimal()
```


## USArrests Plot

Produce a ggplot from the USArrests dataset from the *datasets* package. Place *Assault* on the x-axis, *Murder* on the y, and the points are coloured by *UrbanPop*. Choose a divergeing colour scale using *scale_color_gradient2()*. Pick contrasting end points using *scale_color_gradient2(low = "some-colour", high = "some-different-colour", mid = 66)*. Note the *mid = 66* sets the midpoint of *UrbanPop*, this is the median % for the 50 states.

```{r arrests, exercise=TRUE, exercise.eval = TRUE}

```


```{r arrests-hint}
USArrests %>% 
  ggplot(aes(Assault, Murder, col = UrbanPop)) + 
  geom_point() + 
  scale_colour_gradient2(low = "darkolivegreen", midpoint = 66, high = "orchid") +
  theme_minimal()
```

## Penguin Facets

Draw three side by side histograms of *body_mass_g* from the *penguins* dataset, faceting by *island*.

```{r penguin-histograms, exercise=TRUE, exercise.eval = TRUE}

```

```{r penguin-histograms-hint}
palmerpenguins::penguins %>% 
  ggplot(aes(body_mass_g)) + 
  geom_histogram() + 
  facet_wrap(~island) +
  theme_minimal()
```

## Gapminder Facets

Produce the plot below of data from the gapminder dataset. It shows population (called *pop* in the dataset) over years for different countries, faceted by *continent*. The light blue lines are individual countries, the dark blue line is the fitted line for the continent. Hints: use geom_line(aes(group = country), col = "lightblue") for the country lines, geom_smooth(se = F) for the fitted lines, and a facet_grid() with five columns. Include *scale_y_log10(label = scales::comma_format())* to get a nice shape on the y axis.

```{r gapminder-plot, echo = F}
gapminder::gapminder %>% 
  ggplot(aes(year, pop)) + 
  geom_line(aes(group = country), col = "lightblue") + 
  geom_smooth(se = F) + 
  facet_wrap(~continent, ncol = 5) + 
  scale_y_log10(label = scales::comma_format()) + 
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 20, family = "Ink Free"),
        axis.text.x = element_text(size = 10)) + 
  scale_x_continuous(breaks = c(1950, 1970, 1990))
```


```{r gapminder, exercise=TRUE, exercise.eval = FALSE}
gapminder::gapminder %>% 
  ggplot(aes(year, pop)) + 
  geom_line(aes(group = country), col = "lightblue") +
  theme_minimal()
```


```{r gapminder-hint}
gapminder::gapminder %>% 
  ggplot(aes(year, pop)) + 
  geom_line(aes(group = country), col = "lightblue") + 
  geom_smooth(se = F) + 
  facet_wrap(~continent, ncol = 5) + 
  scale_y_log10(label = scales::comma_format()) + 
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(size = 20, family = "Ink Free"),
        axis.text.x = element_text(size = 10)) + 
  scale_x_continuous(breaks = c(1950, 1970, 1990))
```

